---
title: "Predator-prey model with two interactions"
author: "Ellinor Jakobsson"
date: '2022-11-14'
output: html_document
---
  
#### Tips for working in Rmarkdown: 
* If you haven't worked with RMarkdown before you may need to *install the package* by typing this code *install.packages("rmarkdown")*  
* To close all chunks on windows: *Alt+o*  
* To open all chunks on windows: *Alt+shift+o*  
* To compile the script into an HTML-file, press *knit*  

```{r include=FALSE}
library(deSolve) #Loading all packages
library(ggplot2)
library(tidyr)
```

#### This is the **model** of two interactions  
1. Can you understand what the different parts of the differential equations mean biologically?

```{r echo=TRUE, warning=FALSE}
predpreyLV <- function(t, y, params) {
  R <- y[1]
  C <- y[2]
  with(as.list(params), {
    dR.dt <- r * R * (1 - R / k) - a * R * C #Building the model
    dC.dt <- e * a * R * C - d * C
    return(list(c(dR.dt, dC.dt))) #list of the differential equations, add
  })
} 
```

#### These are the **parameters** specified
2. Discuss and *understand* what role they play in the equation  
```{r}
r <- 2 #Intrinsic growth rate of resource (logistic)
a <- 0.2 #Attack rate of predator on resource
e <- 0.2 #Efficacy of take-up from resource to predator
d <- 0.4 #Death rate of predator
k <- 50 #Carrying capacity of resource

R0 <- 2 #Initial resource population
C0 <- 6 #Initial predator population
params1 <- c(r = r, a = a, e = e, d = d, k = k) # specify parameters
```

```{r warning=FALSE, include=FALSE}
####RUN MODEL####
MaxTime <- 50 # length of the time series, can be changed at any time
Time <- seq(0, MaxTime, by = 0.1) # here we produce a sequence of time points at which we will determine the population densities
LV.out1 <- ode(c(R0, C0), Time, predpreyLV, params1) # here we call the function "ode" that does the hard work of numerical integration and which is part of the package "deSolve"
```

#### **Plot** the model...
* ...using *matplot*  
```{r echo=FALSE}
matplot(Time, (LV.out1[, 2:3]), type = "l", lty = 1:5, col=1:5, ylab = "Population Density", xlab="Time")
legend("top", c(expression("resource"), expression("consumer")), lty = 1:5, col = 1:5, lwd = 2, bty = "n", cex = 0.75) 

```

* ...using *ggplot*  
```{r echo=FALSE}
#Reshape to long format
LV.out <- as.data.frame(LV.out1)
LV.out$time <- as.numeric(LV.out$time)
LV.out <- gather(LV.out, key = "Trophic level", value = "Population_size", -time)
#Plot
ggplot(data = LV.out, aes(x = time, y = Population_size, col = `Trophic level`)) + geom_line() +
  theme_bw() + 
  labs(x = "Time", y = "Population density") + scale_color_brewer(palette="Set1", labels=c('Resource', 'Consumer'))

```

#### Your **task**    
3. Discuss and interpret how the species interact in the different models and compare them. All the tasks and questions are specified in the document   
4. Hand in one HTML Rmarkdown file with the **results** and **conclusions** of all the models you have built during this exercise according to the document     




